---
title: "07_prone_icd"
author: "Cathy Gao and Chad Hochberg"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
# Detach tidytable if it is loaded
if ("package:tidytable" %in% search()) {
  detach("package:tidytable", unload = TRUE)
  cat("tidytable has been detached\n")
} else {
  cat("tidytable was not loaded\n")
}

packages <- c("lubridate",
              "tidyverse",
              "dplyr",
              "ggplot2",
              "tidyr",
              "yaml",
              "rprojroot",
              "tableone",
              "readxl",
              "arrow",
              "collapse",
              "data.table",
              "viridis",
              "sandwich", #Cluster Robust Standard Errors
              "lmtest",  #Allows Calculation of Cluster Robust Standard Errors
              "marginaleffects") #Predictions and Average Slopes for Average Marginal Effects

install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}

sapply(packages, install_if_missing)
rm(packages, install_if_missing)

#Set Working Director as the Project Root
project_root <- file.path( find_root(rprojroot::has_dir("CLIF_Proning_Incidence_Severe_ARF")),
  "CLIF_Proning_Incidence_Severe_ARF")
#project_root <- file.path( find_root(rprojroot::has_dir("CLIF_Proning_Incidence_Severe_ARF-main-Jan 2026 rerun")),
#  "CLIF_Proning_Incidence_Severe_ARF-main-Jan 2026 rerun")
knitr::opts_knit$set(root.dir = project_root)
getwd()


#Use Dplyr select as default
select <- dplyr::select

#Load Config and Setting
config = yaml::read_yaml(paste0(project_root, '/config/config.yaml'))

### user entered site details ---------------------------------------------------
site             = tolower(config$site)
file_type        = tolower(config$file_type)
tables_location  = config$clif_data_path
project_location = config$project_location
site_time_zone = config$site_time_zone

setwd(project_location)
```

```{r Analysis Parameters}
### Hardcoded Analysis Parameters -----------------------------------------------

# Cohort Restriction: Only Include Patients Enrolled in This Calendar Year or Later
# (ICD-10-PCS Procedural Coding for Proning is Assumed More Complete/Reliable From This Point On)
cohort_year_cutoff <- 2023

# ICD-10-PCS Codes for Prone-Positioning Procedures, by Duration
proning_codes <- c(
  "5A09B5K",
  "5A09C5K",
  "5A09D5K"
)

proning_code_map <- c(
  "5A09B5K" = "Assistance with Resp Ventilation, <8 Hrs, Intub Prone",
  "5A09C5K" = "Assistance with Resp Ventilation, 8-24 Hrs, Intub Prone",
  "5A09D5K" = "Assistance with Resp Ventilation, >24 Hrs, Intub Prone"
)

proning_pattern <- paste0("^(", paste(proning_codes, collapse = "|"), ")")

# ICD-10-PCS Codes for Mechanical Ventilation Duration
ventilation_codes <- c(
  "5A1935Z",
  "5A1945Z",
  "5A1955Z"
)

ventilation_code_map <- c(
  "5A1935Z" = "Respiratory Ventilation, Less Than 24 Consecutive Hours",
  "5A1945Z" = "Respiratory Ventilation, 24-96 Consecutive Hours",
  "5A1955Z" = "Respiratory Ventilation, Greater Than 96 Consecutive Hours"
)

ventilation_pattern <- paste0("^(", paste(ventilation_codes, collapse = "|"), ")")
```

```{r Input and Output File Paths}
### Input Data Paths -------------------------------------------------------------
path_prone_analytic_data <- paste0(project_location, '/project_tables/prone_analytic_data.parquet')
path_cohort_hosp_ids     <- paste0(project_location, '/project_tables/cohort_hospitalization_ids.RData')
path_patient_procedures  <- paste0(tables_location, '/clif_patient_procedures.parquet')
path_hospital_diagnosis  <- paste0(tables_location, '/clif_hospital_diagnosis.parquet')

### Output Directories ------------------------------------------------------------
dir_project_output <- paste0(project_location, "/project_output")
icd_output_dir      <- paste0(dir_project_output, "/icd_revision_results_UPLOAD")

### Output File Paths --------------------------------------------------------------
path_out_summary           <- paste0(icd_output_dir, '/', site, '_summary.csv')
path_out_proning_counts    <- paste0(icd_output_dir, '/', site, '_proning_code_counts.csv')
path_out_vent_counts       <- paste0(icd_output_dir, '/', site, '_vent_code_counts.csv')
path_out_proned_cf         <- paste0(icd_output_dir, '/', site, '_proned_cf.csv')
path_out_vent_cf           <- paste0(icd_output_dir, '/', site, '_vent_cf.csv')
path_out_tableone_cont     <- paste0(icd_output_dir, '/', site, '_tableone_cont.csv')
path_out_tableone_cat      <- paste0(icd_output_dir, '/', site, '_tableone_cat.csv')
path_out_proning_quarter   <- paste0(icd_output_dir, '/', site, '_proning_quarter.csv')
path_out_icd_only_summary  <- paste0(icd_output_dir, '/', site, '_icd_only_summary.csv')

### Plot Output Paths (PNG) --------------------------------------------------------
path_out_proned_cf_plot        <- paste0(icd_output_dir, '/', site, '_proned_cf.png')
path_out_vent_cf_plot          <- paste0(icd_output_dir, '/', site, '_vent_cf.png')
path_out_proning_quarter_plot  <- paste0(icd_output_dir, '/', site, '_proning_quarter.png')
```


```{r Optimize Use of Cores - Code from Pat Lyons}
### threads and ram ------------------------------------------------------------

os_type   = Sys.info()[["sysname"]]
all_cores = parallel::detectCores(logical = TRUE)
all_cores = if (is.na(all_cores)) 1L else as.integer(all_cores)

get_ram_gb = function() {
  tryCatch({
    if (os_type == "Darwin") {
      # macOS: sysctl reports total RAM in bytes
      bytes = suppressWarnings(
        as.numeric(system("sysctl -n hw.memsize", intern = TRUE))
      )
      if (length(bytes) > 0 && !is.na(bytes)) bytes / 1024^3 else NA_real_
    } else {
      # Windows & Linux: ps is usually reliable
      mem_info <- ps::ps_system_memory()
      key <- intersect(c("available", "avail"), names(mem_info))
      val <- if (length(key) > 0) mem_info[[key]] / 1024^3 else NA_real_

      if (is.finite(val)) val else {
        # Linux fallback: read /proc/meminfo directly
        if (file.exists("/proc/meminfo")) {
          kb = suppressWarnings(
            as.numeric(system("awk '/MemAvailable/ {print $2}' /proc/meminfo", intern = TRUE))
          )
          if (length(kb) > 0 && !is.na(kb)) kb / 1024^2 else NA_real_
        } else {
          NA_real_
        }
      }
    }
  }, error = function(e) NA_real_)
}

avail_ram_gb = get_ram_gb()

## choose threads conservatively (00 is light; keep headroom) ------------------

reserve_cores   = 1L
gb_per_thread   = 0.50
max_by_cores    = max(1L, all_cores - reserve_cores)
max_by_memory   = if (is.finite(avail_ram_gb)) max(1L, floor(avail_ram_gb / gb_per_thread)) else max_by_cores
n_threads       = as.integer(max(1L, min(max_by_cores, max_by_memory, 8L)))
n_math_threads  = as.integer(max(1L, min(n_threads, 8L)))

## apply thread settings -------------------------------------------------------

data.table::setDTthreads(threads = n_threads)
collapse::set_collapse(nthreads  = n_threads)
options(arrow.use_threads        = TRUE)
Sys.setenv(ARROW_NUM_THREADS     = n_threads)
options(mc.cores                 = n_threads)

# concise summary
message(
  sprintf("Env OK | OS=%s | Cores=%d | Threads=%d | MathThreads=%d | Avail RAM≈%s GB",
          os_type, all_cores, n_threads, n_math_threads,
          ifelse(is.finite(avail_ram_gb), round(avail_ram_gb, 1), "NA"))
)
```


```{r Create ICD sub Folders}
### file locations -------------------------------------------------------------
if (!dir.exists(dir_project_output)) {
  dir.create(dir_project_output)
}
if (!dir.exists(icd_output_dir)) {
  dir.create(icd_output_dir)
}
```

```{r Function to Convert Datetimes to Local Time Zone}
#Write Function to Convert Datetime Columns to 'UTC'
fn.timezone_tolocal <- function(df) {
  # Get the dataset schema
  sch <- df$schema
  #Which Fields are Datetime Types?
  datetime_fields <- keep(sch$fields, function(field) {
    grepl("^timestamp", field$type$ToString())
  })
  # Extract the names of these datetime columns
  datetime_cols <- map_chr(datetime_fields, "name")

  #Now Convert DateTime Columns to Local Timezone
  df <- df %>%
    mutate(dplyr::across(dplyr::all_of(datetime_cols), ~ cast(.x, arrow::timestamp("us", site_time_zone)))) |>
    compute()
}
```

```{r Function to Decompose Table 1 Summaries for Easy Pooling}
tab_decomp <- function(table) {

  results_list <- list()  # Prepare an empty list to store results


#First Do Continuous Table
if (!is.null(table$ContTable)) {
  strata <- rownames(table$ContTable)
  for (i in strata) {
    # Extract the relevant names and data for the current strata
    rows <- rownames(table$ContTable[[i]])
    cols <- colnames(table$ContTable[[i]])

    x <- as_tibble(table$ContTable[[i]])
    colnames(x) <- cols

    # Create a new tibble with the desired structure
    x <- x |>
      mutate(
        strata = i,
        variable = rows) |>
      relocate(strata, variable, .before = 1)

    results_list[[i]] <- x  # Append result to the list
  }
  # Combine all results into a single tibble
  cont_table <- rowbind(results_list)

} else {
  cat('\n No Continuous Table for These Variables\n')
  cont_table <- NULL  # Assign NULL if there are no continuous tables
}
  #Now Categorical Table
 results_list <- list() #Reinitialize List
if (!is.null(table$CatTable)) {
    strata <- rownames(table$CatTable)
  for (i in strata) {
    category <- table$MetaData$varFactors
    for (z in category) {
    # Extract the relevant names and data for the current strata
    rows <- rownames(table$CatTable[[i]][[z]])
    cols <- colnames(table$CatTable[[i]][[z]])

    x <- as_tibble(table$CatTable[[i]][[z]])
    colnames(x) <- cols

    # Create a new tibble with the desired structure
    x <- x |>
      mutate(
        strata = i,
        variable = z) |>
      relocate(strata, variable, level, .before = 1)

    results_list[[length(results_list) + 1]] <- x  # Append result to the list
    }
      }
  # Combine all results into a single tibble
  cat_table <- rowbind(results_list)
} else {
  cat('\n No Categorical Table for These Variables\n')
  cat_table <- NULL  # Assign NULL if there are no categorical tables
}
  #Put Two Tables into  a List
  table_list <- list(continuous = cont_table, categorical = cat_table)
  return(table_list)
  }
```


```{r Load Site-specific analytic data set and Convert to Local Time Zone}
#NOTE: This Table is Created by the 00_file if you need to regenerate
prone_analytic_df <- open_dataset(path_prone_analytic_data)

#Apply Function to Convert to Local Timezone
prone_analytic_df <- fn.timezone_tolocal(prone_analytic_df)

#Bring Into R Environment as Dataframe
prone_analytic_df <- prone_analytic_df |> collect()

#Limit to Cohort Year Cutoff
prone_analytic_df <- prone_analytic_df |>
  filter(year(t_enrollment)>=cohort_year_cutoff) |>
  #Make Female Variable
  mutate(female=fifelse(tolower(sex_category)=='female', 1, 0))
```

#Derive Covariates Needed for Table 1 That Are Not Persisted in prone_analytic_data.parquet
#Same Logic as Notebooks 01/02/06 (each independently re-derives these from prone_analytic_data since none of them write them back to disk)
#NOTE: We Intentionally Do NOT Re-Derive 'study_period'/'period_sarscov2' Here - Since This Cohort is Already Filtered to year(t_enrollment)>=cohort_year_cutoff,
#Every Row Falls in the 'Post-COVID' Period (Study Month >=51 i.e. March 2022 Onward) in Notebooks 01/02/06's Period Definition,
#So Those Columns Would Be Constant/Non-Informative For This Table 1 and Are Left Out of 'to_tab'/'factors_tab' Below
```{r Derive Table 1 Covariates (year, admit_to_enrolled, ett_to_enrolled, severe_ards)}
prone_analytic_df <- prone_analytic_df |>
  #Calendar Year of Enrollment (Same Logic as Notebook 01's 'Define Study Months/Quarters' Chunk)
  mutate(year=year(t_enrollment)) |>
  #Timing Covariates (Same Logic as Notebook 01)
  mutate(admit_to_enrolled=as.duration(t_enrollment-final_admission_dttm)/dhours(1)) |>
  mutate(ett_to_enrolled=as.duration(t_enrollment-vent_episode_start)/dhours(1)) |>
  #Severe ARDS Flag (Same Logic as Notebook 01/02/06: Min P/F Ratio <100)
  mutate(severe_ards=fifelse(min_pf_ratio<100, 1, 0))
```

```{r}
#Get list of Hospitalization IDs; some admissions (defined as patient_id+encounter_block, can span multiple hospitalization_ids [i.e. transferred between site hospitals])

load(path_cohort_hosp_ids)

cohort_hospitalization_ids <- cohort_hospitalization_ids |>
  mutate(hospitalization_id = as.character(hospitalization_id))

# QC: A Duplicate hospitalization_id Here Would Fan Out in the proning_icd/ventilation_icd Joins
# Below, Risking One Patient's ICD Billing Codes Getting Misattributed to Another
dup_hospitalization_ids <- cohort_hospitalization_ids |>
  count(hospitalization_id) |>
  filter(n > 1)

if (nrow(dup_hospitalization_ids) > 0) {
  warning(
    nrow(dup_hospitalization_ids), " duplicate hospitalization_id(s) found in cohort_hospitalization_ids at site '",
    site, "'. This can cause ICD billing codes to be misattributed across patients in the proning_icd/",
    "ventilation_icd joins downstream. Investigate before trusting this site's results."
  )
}
```


```{r Raw proning code counts}
#  counts by code and name
proning_counts_proc <- open_dataset(path_patient_procedures) |>
  collect() |>
  filter(
    toupper(procedure_code_format) %in% c("ICD10PCS", "ICD-10-PCS"),
    grepl(proning_pattern, procedure_code, ignore.case = TRUE)
  ) |>
  mutate(
    procedure_code = toupper(procedure_code),
    procedure_name = proning_code_map[procedure_code]
  ) |>
  count(procedure_code, procedure_name, sort = TRUE)
```

```{r Raw proning code counts check diagnosis table}
# optional counts by code and name
proning_counts_dx <- open_dataset(path_hospital_diagnosis) |>
  collect() |>
  filter(
    grepl(proning_pattern, diagnosis_code, ignore.case = TRUE)
  ) |>
  mutate(
    diagnosis_code = toupper(diagnosis_code),
    procedure_name = proning_code_map[diagnosis_code]
  ) |>
  count(diagnosis_code, procedure_name, sort = TRUE)
```


```{r See which table is bigger}
# QC: If Neither Raw Table Has Any Matching Proning Code at All, earliest_procedure_billed_dttm
# Below Would Resolve to NA, Which Would Silently Filter the Entire Cohort to Zero Rows Downstream
# With No Explanation. Fail Fast Here Instead.
if (nrow(proning_counts_proc) == 0 && nrow(proning_counts_dx) == 0) {
  stop(
    "No ICD-10-PCS proning codes (", paste(proning_codes, collapse = ", "), ") found in either ",
    "clif_patient_procedures or clif_hospital_diagnosis at site '", site, "'. This ICD validation ",
    "sub-study cannot run for this site without at least one matching billing code."
  )
}

# Tie-Break Rule: If Both Tables Have the Same Number of Matching Rows, Prefer clif_patient_procedures
larger_proning_table <- dplyr::case_when(
  nrow(proning_counts_dx) > nrow(proning_counts_proc) ~ "clif_hospital_diagnosis",
  TRUE ~ "clif_patient_procedures"
)

# QC: All Downstream Code (Earliest-Billing-Date, proning_icd, ventilation_icd) Assumes the
# clif_patient_procedures Schema (procedure_code_format/procedure_code/procedure_billed_dttm).
# clif_hospital_diagnosis Does Not Share This Schema (It Only Has diagnosis_code), So Fail Fast
# Here With a Clear Message Instead of Letting It Crash Several Chunks Downstream.
if (larger_proning_table != "clif_patient_procedures") {
  stop(
    "larger_proning_table resolved to '", larger_proning_table, "' at site '", site, "'. ",
    "Downstream code (earliest-billing-date, proning_icd, ventilation_icd chunks) assumes the ",
    "clif_patient_procedures schema and does not support clif_hospital_diagnosis. ",
    "This notebook needs to be updated to handle that schema before proceeding at this site."
  )
}

# Path to Whichever Raw Table Was Selected Above (Reused for All Subsequent ICD-Code Extraction)
path_larger_proning_table <- paste0(tables_location, "/", larger_proning_table, ".parquet")
```



```{r}
# earliest documented instance by code and name

earliest_procedure_billed_dttm <- open_dataset(path_larger_proning_table) |>
  collect() |>
  filter(
    toupper(procedure_code_format) %in% c("ICD10PCS", "ICD-10-PCS"),
    grepl(proning_pattern, procedure_code, ignore.case = TRUE)
  ) |>
  summarise(
    earliest_procedure_billed_dttm = min(procedure_billed_dttm, na.rm = TRUE)
  ) |>
  pull(earliest_procedure_billed_dttm) |>
  with_tz(tzone = site_time_zone)
```


```{r}
# Prepare a data frame with the table name and earliest procedure
proning_summary <- data.frame(
  proning_table = larger_proning_table,
  earliest_procedure_billed_dttm = earliest_procedure_billed_dttm
)
write_csv(proning_summary, path_out_summary)
```


```{r Filter Dataset Only to After The First Instance of Billing}
prone_analytic_df <- prone_analytic_df |>
  filter(t_enrollment>=earliest_procedure_billed_dttm)

cohort_hospitalization_ids <- cohort_hospitalization_ids |>
  semi_join(prone_analytic_df, by = c('patient_id', 'encounter_block'))
```




```{r}
proning_icd <- open_dataset(path_larger_proning_table) |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  dplyr::semi_join(cohort_hospitalization_ids, by = "hospitalization_id") |>
  dplyr::left_join(
    cohort_hospitalization_ids |>
      select(hospitalization_id, patient_id, encounter_block),
    by = "hospitalization_id"
  ) |>
  collect() |>
  filter(
    toupper(procedure_code_format) %in% c("ICD10PCS", "ICD-10-PCS"),
    grepl(proning_pattern, procedure_code, ignore.case = TRUE)
  ) |>
  mutate(
    procedure_code = toupper(procedure_code)
  ) |>
  group_by(patient_id, encounter_block) |>
  summarise(
    proning_lt8h   = as.integer(any(procedure_code == "5A09B5K")),
    proning_8_24h  = as.integer(any(procedure_code == "5A09C5K")),
    proning_gt24h  = as.integer(any(procedure_code == "5A09D5K")),
    .groups = "drop"
  )
```

```{r}
ventilation_icd <- open_dataset(path_larger_proning_table)|>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  dplyr::semi_join(cohort_hospitalization_ids, by = "hospitalization_id") |>
  dplyr::left_join(
    cohort_hospitalization_ids |>
      select(hospitalization_id, patient_id, encounter_block),
    by = "hospitalization_id"
  ) |>
  collect() |>
  filter(
    toupper(procedure_code_format) %in% c("ICD10PCS", "ICD-10-PCS"),
    grepl(ventilation_pattern, procedure_code, ignore.case = TRUE)
  ) |>
  mutate(
    procedure_code = toupper(procedure_code)
  ) |>
  group_by(patient_id, encounter_block) |>
  summarise(
    vent_lt24h   = as.integer(any(procedure_code == "5A1935Z")),
    vent_24_96h  = as.integer(any(procedure_code == "5A1945Z")),
    vent_gt96h   = as.integer(any(procedure_code == "5A1955Z")),
    .groups = "drop"
  )
```



```{r}
joined_df <- dplyr::left_join(
  prone_analytic_df,
  proning_icd,
  by = c("patient_id", "encounter_block")
)

joined_df <- dplyr::left_join(
  joined_df,
  ventilation_icd,
  by = c("patient_id", "encounter_block")
)

# QC: proning_icd/ventilation_icd Are Grouped to One Row per patient_id x encounter_block, So These
# left_joins Should Not Change Row Count. A Mismatch Here Would Indicate Unexpected Row Duplication.
if (nrow(joined_df) != nrow(prone_analytic_df)) {
  warning(
    "joined_df has ", nrow(joined_df), " rows but prone_analytic_df has ", nrow(prone_analytic_df),
    " rows at site '", site, "'. Expected the proning_icd/ventilation_icd joins to be 1:1 ",
    "- investigate for unexpected row duplication."
  )
}
```


```{r}
proning_code_counts <- joined_df %>%
  #Total Number of Hospitals
  mutate(
    N_hospitals = length(unique(joined_df$hospital_id)),
    #N Hospitals with Proning ICD
    n_hospitals_proneICD = length(unique(joined_df$hospital_id[!is.na(joined_df$proning_lt8h) | !is.na(joined_df$proning_8_24h) | !is.na(joined_df$proning_gt24h)]))
  ) |>
  group_by(proning_lt8h, proning_8_24h, proning_gt24h) %>%
  summarise(
    count = n(),
    N_hospitals_site=median(N_hospitals),
    n_hospitals_with_prone_icd=median(n_hospitals_proneICD, na.rm=T),
    median_pt_prone_duration = median(median_pt_prone_duration, na.rm = TRUE),
    .groups = "drop"
  )


write_csv(proning_code_counts, path_out_proning_counts)


vent_code_counts <- joined_df %>%
  group_by(vent_lt24h, vent_24_96h, vent_gt96h) %>%
  summarise(
    count = n(),
    median_vent_duration_hours = median(vent_duration_hours, na.rm = TRUE),
    .groups = "drop"
  )

write_csv(vent_code_counts, path_out_vent_counts)
```

```{r}
joined_df <- joined_df %>%
  mutate(any_proning_icd = as.integer(proning_lt8h == 1 | proning_8_24h == 1 | proning_gt24h == 1),
         any_proning_icd = ifelse(is.na(any_proning_icd), 0L, any_proning_icd))

joined_df <- joined_df %>%
  mutate(any_ventilation_icd = as.integer(vent_lt24h == 1 | vent_24_96h == 1 | vent_gt96h == 1),
         any_ventilation_icd = ifelse(is.na(any_ventilation_icd), 0L, any_ventilation_icd))

joined_df$any_ventilation_flowsheet = 1
```


```{r}
# create confusion table
confusion <- joined_df %>%
  summarise(
    TP = sum(any_proning_icd == 1 & proned == 1, na.rm = TRUE),
    TN = sum(any_proning_icd == 0 & proned == 0, na.rm = TRUE),
    FP = sum(any_proning_icd == 1 & proned == 0, na.rm = TRUE),
    FN = sum(any_proning_icd == 0 & proned == 1, na.rm = TRUE)
  )

# calculate sensitivity and specificity
sensitivity <- confusion$TP / (confusion$TP + confusion$FN)
specificity <- confusion$TN / (confusion$TN + confusion$FP)

# QC: A Zero Denominator Here Means No Truly-Proned (or No Truly-Not-Proned) Patients in the Filtered
# Cohort, Which Would Silently Produce NaN Rather Than a Meaningful Sensitivity/Specificity
if ((confusion$TP + confusion$FN) == 0) {
  warning("sensitivity is NaN at site '", site, "': no flowsheet-proned (proned==1) patients in the filtered cohort.")
}
if ((confusion$TN + confusion$FP) == 0) {
  warning("specificity is NaN at site '", site, "': no flowsheet-not-proned (proned==0) patients in the filtered cohort.")
}

sensitivity
specificity

```



```{r}
confusion_df_proning <- joined_df %>%
  count(proned, any_proning_icd) %>%
  rename(Count = n) %>%
  complete(proned = c(0, 1),
           any_proning_icd = c(0, 1),
           fill = list(Count = 0))

write_csv(confusion_df_proning, path_out_proned_cf)

plot_a <- ggplot(confusion_df_proning, aes(x = any_proning_icd, y = proned, fill = Count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Count), color = "blue", size = 6) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_x_continuous(breaks = c(0, 1), labels = c("0", "1")) +
  scale_y_continuous(breaks = c(0, 1), labels = c("0", "1")) +
  labs(
    title = "Proning Confusion Matrix",
    x = "any_proning_icd_fill",
    y = "proned flowsheet"
  ) +
  coord_fixed() +
  theme_minimal(base_size = 10)
plot_a
ggsave(path_out_proned_cf_plot, plot = plot_a, width = 6, height = 6)
```



```{r}
# create confusion matrix table


confusion_df_vent <- joined_df %>%
  count(any_ventilation_flowsheet, any_ventilation_icd) %>%
  rename(Count = n) %>%
  complete(any_ventilation_flowsheet = c(0, 1),
           any_ventilation_icd = c(0, 1),
           fill = list(Count = 0))

write_csv(confusion_df_vent, path_out_vent_cf)


plot_b <- ggplot(confusion_df_vent, aes(x = any_ventilation_icd, y = any_ventilation_flowsheet, fill = Count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Count), color = "white", size = 6) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  scale_x_continuous(breaks = c(0, 1), labels = c("0", "1")) +
  scale_y_continuous(breaks = c(0, 1), labels = c("0", "1")) +
  labs(
    title = "Ventilation Confusion Matrix",
    x = "any_ventilation_icd",
    y = "any_ventilation_flowsheet"
  ) +
  coord_fixed() +
  theme_minimal(base_size = 10)
plot_b
ggsave(path_out_vent_cf_plot, plot = plot_b, width = 6, height = 6)

```



```{r}
joined_df <- joined_df |>
  mutate(
    race_category = case_when(
      tolower(race_category) %in% c("other", "unknown") ~ "Other/Unknown",
      TRUE ~ race_category
    )
  )

# Define variables for TableOne

to_tab <- c(
  'age_at_admission',
  'sex_category',
  'race_category',
  'ethnicity_category',
  'bmi',
  'hospital_id',
  'year',
  'eligible_by_proseva',
  'eligible_by_prone',
  'admit_to_enrolled',
  'ett_to_enrolled',
  'or_before_enrollment',
  'min_pf_ratio',
  'severe_ards',
  'first_proseva_mode',
  'first_proseva_peep',
  'first_proseva_fio2',
  'sofa_score',
  'nee_pressor_dose',
  "prone_episodes",
  "median_pt_prone_duration",
  "mean_pt_prone_duration",
  "vent_duration_hours",
  "death_or_hospice"
)

#Factor Variables
factors_tab <- c(
  'sex_category',
  'race_category',
  'ethnicity_category',
  'hospital_id',
  'year',
  'or_before_enrollment',
  'first_proseva_mode',
  'nee_pressor_dose',
  'severe_ards',
  'eligible_by_proseva',
  'eligible_by_prone',
  "death_or_hospice"
)

# Create TableOne
tableone_cohort <- CreateTableOne(
  data=joined_df,
  vars=to_tab,
  factorVars=factors_tab,
  strata='proned',
  addOverall = T)

# Print TableOne
summary(tableone_cohort)
table_list <- tab_decomp(tableone_cohort)
table_list$continuous <- table_list$continuous |> mutate(strata=fcase(strata=='0', 'not_proned', strata=='1', 'proned', strata=='Overall', 'Overall'))
table_list$categorical <- table_list$categorical |> mutate(strata=fcase(strata=='0', 'not_proned', strata=='1', 'proned', strata=='Overall', 'Overall'))
write_csv(table_list$continuous, path_out_tableone_cont)
write_csv(table_list$categorical, path_out_tableone_cat)

```




```{r}
# Summarize counts per year-quarter
proning_quarter <- joined_df |>
  mutate(
    year = year(final_admission_dttm),
    qtr = quarter(final_admission_dttm),
    year_quarter = paste0(year, "-Q", qtr)
  ) |>
  group_by(year_quarter) |>
  summarise(
    any_proning_icd = sum(any_proning_icd, na.rm = TRUE),
    proned = sum(proned, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(
    cols = c(any_proning_icd, proned),
    names_to = "proning_type",
    values_to = "count"
  )


write_csv(proning_quarter, path_out_proning_quarter)


# Plot bar chart per quarter
plot_c <- ggplot(proning_quarter, aes(x = year_quarter, y = count, fill = proning_type)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("any_proning_icd" = "steelblue", "proned" = "darkorange")) +
  labs(
    title = "Proning Counts per Quarter",
    x = "Year-Quarter",
    y = "Number of Patients",
    fill = "Proning Type"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot_c
ggsave(path_out_proning_quarter_plot, plot = plot_c, width = 10, height = 6)


```


# QC
```{r}
proning_icd_events <- open_dataset(path_patient_procedures) |>
  mutate(hospitalization_id = as.character(hospitalization_id)) |>
  dplyr::semi_join(cohort_hospitalization_ids, by = "hospitalization_id") |>
  dplyr::left_join(
    cohort_hospitalization_ids |>
      select(hospitalization_id, patient_id, encounter_block),
    by = "hospitalization_id"
  ) |>
  collect() |>
  filter(
    toupper(procedure_code_format) %in% c("ICD10PCS", "ICD-10-PCS"),
    grepl(proning_pattern, procedure_code, ignore.case = TRUE)
  ) |>
  mutate(
    procedure_code = toupper(procedure_code)
  ) |>
  transmute(
    patient_id,
    encounter_block,
    icd_prone_dttm = procedure_billed_dttm
  ) |>
  filter(!is.na(icd_prone_dttm))
```

```{r}
first_proning_icd_df <- proning_icd_events %>%
  group_by(patient_id, encounter_block) %>%
  summarise(
    first_proning_icd_dttm = min(icd_prone_dttm),
    .groups = "drop"
  )

joined_df_qc <- joined_df %>%
  left_join(
    first_proning_icd_df,
    by = c("patient_id", "encounter_block")
  )

icd_only_prone_df <- joined_df_qc %>%
  filter(
    any_proning_icd == 1,
    proned == 0
  )

icd_only_prone_qc_table <- icd_only_prone_df %>%
  mutate(
    icd_to_clif_time_diff_hrs = as.numeric(
      difftime(first_prone_time, first_proning_icd_dttm, units = "hours")
    )
  ) %>%
  select(
    patient_id,
    encounter_block,
    vent_duration_hours,
    first_proning_icd_dttm,
    first_prone_time,
    icd_to_clif_time_diff_hrs,
    proning_lt8h,
    proning_8_24h,
    proning_gt24h
  )

icd_only_summary <- icd_only_prone_qc_table %>%
  summarise(
    n_icd_only = n(),
    median_vent_hours = median(vent_duration_hours, na.rm = TRUE),
    median_icd_to_clif_diff_hrs = median(icd_to_clif_time_diff_hrs, na.rm = TRUE),
    pct_short_prone = mean(proning_lt8h == 1, na.rm = TRUE),
    pct_medium_prone = mean(proning_8_24h == 1, na.rm = TRUE),
    pct_long_prone = mean(proning_gt24h == 1, na.rm = TRUE)
  )

write_csv(icd_only_summary, path_out_icd_only_summary)

cat("\nICD-only summary data saved as '",
    path_out_icd_only_summary, "'\n", sep = "")
```
