---
title: "cohort_identification_prone_incidence"
author: "Anna Barker and Chad Hochberg"
date: "`r Sys.Date()`"
output: html_document
---

```{r Load Needed Libraries, include=FALSE}
packages <- c("duckdb", "lubridate", "tidyverse", "dplyr","table1", "broom", "arrow", "rvest", "readr", "fst", "data.table", "collapse")

install_if_missing <- function(package) {
  if (!require(package, character.only = TRUE)) {
    install.packages(package, dependencies = TRUE)
    library(package, character.only = TRUE)
  }
}

sapply(packages, install_if_missing)
rm(packages, install_if_missing)

#Use Dplyr select as default
select <- dplyr::select

```


#Objective: identify a cohort of hospitalizations with patients who receive mechanical ventilation from CLIF tables
Identify patients admitted to the hospital in a given date range. Export a list of `hospitalization_id` and filtered CLIF tables for the identified hospitalizations.

# Specify inpatient cohort parameters
Age >= 18 Years
PF <=150 on PEEP >= 5 and FiO2 >= 0.6
Criteria are met in the first 36 hours of IMV initiation (t_PROSEVA_first) 
AND
Confirmed on second qualifying blood gas in a 12-hour window beginning 12 hours after the first eligible blood gas (t_proseva_first + 12 hours > t_proseva_second < t_proseva_first + 24 hours) 
OR
Patient is proned within 24 hours of initial qualifying blood gas (t_PROSEVA_first < t_proning < t_PROSEVA_first + 24 hours) 

## Date range
Specify the start and end dates for the cohort
```{r Start Times and Whether Pediatric Patients Are Included}
start_date <- "2018-01-01"
end_date <- "2023-12-31"

include_pediatric <- FALSE
include_er_deaths <- TRUE
```

#Specify required CLIF Tables
```{r}
#List of table names from CLIF 2.0
tables <- c("patient", "hospitalization", "vitals", "labs", 
            "medication_admin_continuous", "adt", 
            "patient_assessments", "respiratory_support", "position", 
            "dialysis", "intake_output", "ecmo_mcs", "procedures", 
            "admission_diagnosis", "provider", "sensitivity", 
            "medication_orders", "medication_admin_intermittent", 
            "therapy_details", "microbiology_culture", "sensitivity", "microbiology_nonculture")

# Tables that should be set to TRUE for this project
true_tables <- c("patient", "hospitalization", "adt",
                 "vitals", "labs", "medication_admin_continuous", "respiratory_support",
                 "position")

# Create a named vector and set the boolean values
table_flags <- setNames(tables %in% true_tables, tables)
```

#Specify File Paths, Project Path and Site - SITE SPECIFIC
#Load the Required CLIF Tables (or if using ARROW/parquet refer to table in situ on storage drive)
```{r}
tables_location <- '~/workspace/Storage/chochbe1/JH_CCRD/CLIF/rclif'
project_location <- '~/workspace/Storage/chochbe1/JH_CCRD/CLIF/CLIF_Projects/CLIF_prone_incidence'
site <- "Hopkins"
file_type <- 'parquet'

#Create Sub Folders within Project Folder
# Check if the output directory exists; if not, create it
if (!dir.exists(paste0(project_location, "/project_tables"))) {
  dir.create(paste0(project_location, "/project_tables"))
}
if (!dir.exists(paste0(project_location, "/project_output"))) {
  dir.create(paste0(project_location, "/project_output"))
}
```

```{r}
# List all CLIF files in the directory
clif_table_filenames <- list.files(path = tables_location, 
                                   pattern = paste0("^clif_.*\\.", file_type, "$"), 
                                   full.names = TRUE)

# Extract the base names of the files (without extension)
clif_table_basenames <- basename(clif_table_filenames) |>
  str_remove(paste0("\\.", file_type, "$"))

# Create a lookup table for required files based on table_flags
required_files <- paste0("clif_", names(table_flags)[table_flags])

# Check if all required files are present
missing_tables <- setdiff(required_files, clif_table_basenames)
if (length(missing_tables) > 0) {
  stop(paste("Error: Missing required tables:", paste(missing_tables, collapse = ", ")))
}

# Filter only the filenames that are required
required_filenames <- clif_table_filenames[clif_table_basenames %in% required_files]

# Read the required files into a list of data frames
if (file_type == "parquet") {
  data_list <- lapply(required_filenames, open_dataset)
} else if (file_type == "csv") {
  data_list <- lapply(required_filenames, read_csv)
} else if (file_type == "fst") {
  data_list <- lapply(required_filenames, read.fst)
} else {
  stop("Unsupported file format")
}

# Assign the data frames to variables based on their file names
for (i in seq_along(required_filenames)) {
  # Extract the base name of the file (without extension)
  object_name <- str_remove(basename(required_filenames[i]), paste0("\\.", file_type, "$"))
  # Make the object name valid for R (replace invalid characters with underscores)
  object_name <- make.names(object_name)
  # Assign the tibble to a variable with the name of the file
  assign(object_name, data_list[[i]])
}
```

#Now Ready to Identify Cohort
#Identify Hospitalizations for Adults >= 18 Who Were Ever in an ED, Ward or ICU
#At This Stage Identify Encounters that Are 'Linked', that is continuous admission within the health system but in different hospitals (or sometimes ED to inpatient in same hospital)
#Script below will create an 'encounter_block' variable toto identify admissions at the patient_id level that are linked
```{r Identify Hospitalizations in Right Time Frame and Age}
clif_hospitalization <- clif_hospitalization |>
   filter(admission_dttm >= as.POSIXct(start_date, tz = "UTC") &
   admission_dttm <= as.POSIXct(end_date, tz = "UTC")) |>
   compute()

if (!include_pediatric) {
  clif_hospitalization <- clif_hospitalization |>
    filter(age_at_admission >= 18) |>
    compute()
}

#To be Thorough Will Keep WARD, ICU and ER Encounters (in case of ED to ICU within one system)
inpatient_hospitalization_ids <- clif_adt |>
  filter(location_category %in% c("Ward", "ICU", "ER")) |>
  select(hospitalization_id) |>
  collect()

clif_hospitalization <- clif_hospitalization |>
  collect() |> #Have to Bring in to Environment To Correctly Filter IDs (at JHH)
  filter(hospitalization_id %in% inpatient_hospitalization_ids$hospitalization_id) |>
  as_arrow_table() #Put Back to Arrow Table
rm(inpatient_hospitalization_ids)

#Create an Hospital Block ID - This is to Identify Continuous Hospitalizations When Patients Are Transferred Between Hospitals in One Health System
#This code is intended be robust to various ways encounters may be coded in CLIF databases
hospital_blocks <- clif_hospitalization |>
  select(patient_id, hospitalization_id, admission_dttm, discharge_dttm) |>
  arrange(patient_id, admission_dttm) |>
  collect()

#Identify Admissions That Occur Within 3 Hours of a Discharge (Will Consider Those Linked and as Part of One Continuous Encounter)
#Use Data Table for Speed
linked_encounters <- setDT(hospital_blocks)
#Create a Variable for the time of the next admission and time of previous discharge
linked_encounters[, ':=' (next_admit_dttm = data.table::shift(admission_dttm, n=1, type = "lead")), by = patient_id]
linked_encounters[, ':=' (prev_dc_dttm = data.table::shift(discharge_dttm, n=1, type = "lag")), by = patient_id]
#Calculates Time Between Discharge and Next Admit
linked_encounters[, next_diff_time := difftime(next_admit_dttm, discharge_dttm, units = "hours")]
linked_encounters[, prev_diff_time := difftime(admission_dttm, prev_dc_dttm, units = "hours")]

#Now Create Variable Indicating a Linked Encounter (next_admit-dc time <6 hours or prev_dc-admint <6 hours)
linked_encounters[, linked := fcase(
  (next_diff_time <6 | prev_diff_time <6), 1)]
#Filter to Only Linked Encounters and number them
linked_encounters <- linked_encounters[linked==1]
#This Identifies the First Encounter in a Series of Linked Encounters
linked_encounters[, first_link := fcase(
  (rowid(linked)==1 | (next_diff_time<6 & prev_diff_time>6)), 1
), by = patient_id]
#Now Numbers Encounters, easier in dplyr
#Filter to Just First Links, Number them and then Remerge with linked encounters
temp <- as_tibble(linked_encounters) |>
 filter(first_link==1) |>
 group_by(patient_id) |>
 mutate(link_group=row_number()) |>
 ungroup() |>
 select(hospitalization_id, link_group) 
linked_encounters <- as_tibble(left_join(linked_encounters, temp)) |>
  fill(link_group, .direction = c("down")) |>
  #Create a Variable Indicating Which Number of LIinked Encounter the Encounter is
  group_by(patient_id, link_group) |>
  mutate(link_number=row_number()) |>
  ungroup() |>
  select(hospitalization_id, linked, link_number)
rm(temp)

#Now Join Back to Hospitalization Table
clif_hospitalization <- clif_hospitalization |>
  left_join(linked_encounters) |>
  mutate(linked=if_else(is.na(linked), 0, linked)) |>
  compute()

#Pull Out the Any Linked Encounter that Is NOt the First Encounter and Assign Each Encounter an Encounter Block ID in the Original clif_hospitalization table
df_link <- clif_hospitalization |>
  filter(link_number>1) |>
  collect()

clif_hospitalization <- clif_hospitalization |>
  group_by(patient_id) |>
  arrange(patient_id, admission_dttm) |>
  #Remove Link Numbers that Are Not First in Link Encounter
  filter(link_number==1 | is.na(link_number)) |>
  #Make Encounter Blocks
  collect() |>
  mutate(encounter_block=row_number()) |>
  rowbind(df_link, fill = TRUE) |> #Bring Back in Link Numbers >1
  group_by(patient_id) |> arrange(patient_id, admission_dttm) |>
  fill(encounter_block, .direction = "down") |>
  ungroup()|>
  #Finally, for Linked Encounters Identify 'Final_admit_date' and 'final_dc_date' which are the first and last dates of a link block
  group_by(patient_id, encounter_block) |>
  mutate(final_admission_dttm=fcase(
    row_number()==1, admission_dttm
  )) |>
  mutate(final_discharge_dttm=fcase(
    row_number()==n(), discharge_dttm
  )) |>
  fill(final_admission_dttm, final_discharge_dttm, .direction = 'updown') |>
  relocate(encounter_block, .after = 'hospitalization_id') |>
  as_arrow_table()

rm(linked_encounters, df_link, hospital_blocks)

#Keep Track for Consort Diagram
patients <- length(unique(clif_hospitalization$patient_id))
encounters <- length(clif_hospitalization$hospitalization_id)
cat('\n In', site, 'CLIF data there are', patients,'unique patients with', encounters, 'encounters \n')
```

#Filter Only to Patients Who Receive Mechanical Ventilation
```{r}
#Bring in Temporary File with patient_id, hospitalization_id and encounter_block
temp_ids <- clif_hospitalization |>
  select(patient_id, hospitalization_id, encounter_block) |>
  collect()

#Identify Patients Who EVER Received Mechanical Ventilation During a Hospitalization
vent <- clif_respiratory_support |>
  #Only Need IDs in the Current Working ClIF hospitalization Table
  filter(hospitalization_id %in% temp_ids$hospitalization_id) |>
  #Will Also Merge in Encounter Block Here - Will Allow Us to Keep Track of LInked Encounters
  left_join(temp_ids) |>
  compute()
rm(temp_ids)

#Identify Those Who Have Ever Been on a Vent During an Encounter Block
vent <- vent |> 
  mutate(on_vent=if_else(device_category=='IMV', 1, 0)) |>
  group_by(patient_id, encounter_block) |>
  mutate(ever_vent=if_else(max(on_vent, na.rm=T)==1, 1, 0)) |>
  filter(ever_vent==1) |>
  select(-ever_vent) |>
  ungroup() |>
  compute()

#Identify First Vent Start Time and Vent Duration
#Keep Track for Consort Diagram
patients <- length(unique(vent$patient_id))
encounters <- length(unique(vent$hospitalization_id))
cat('\n Paitents Receiving Mechanical Ventilation: \n  In', site, 'CLIF data there are', patients,'unique patients with', encounters, 'encounters \n')
```

#Identify Ventilator Episodes
```{r}
#Now Identify Ventilator Episodes and the Duration of Each Episodes
#This Uses Logic Created by Nick Ingraham to Carry Forward Device Category and Device Names
vent <- vent |>
  mutate(
    device_category = 
    if_else(
        is.na(device_category) & is.na(device_name) &
          str_detect(mode_category, 
                     "Pressure Control|Assist Control-Volume Control|Pressure Support/CPAP|Pressure-Regulated Volume Control|SIMV"),
        "IMV",
        device_category
      ),
    device_name = 
      if_else(
        str_detect(device_category, "IMV") & is.na(device_name) &
          str_detect(mode_category, "Pressure Control|Assist Control-Volume Control|Pressure Support/CPAP|Pressure-Regulated Volume Control|SIMV"),
        "IMV",
        device_name
      ),
  ) |>
  collect() |>
  #     If device before is VENT + normal vent things ... its VENT too 
  group_by(patient_id, encounter_block) |>
  arrange(patient_id, recorded_dttm) |>
  mutate(device_category = fifelse(is.na(device_category) & 
                                     lag(device_category == "IMV") & 
                                     tidal_volume_set > 1 & 
                                     resp_rate_set > 1 & 
                                     peep_set > 1, 
                                   "IMV", 
                                   device_category)) |>
  # If device after is VENT + normal vent things ... its VENT too 
  mutate(device_category = fifelse(is.na(device_category) & 
                                     lead(device_category == "IMV") & 
                                     tidal_volume_set > 1 & 
                                     resp_rate_set > 1 & 
                                     peep_set > 1, 
                                   "IMV", 
                                   device_category)) |>
  # doing this for BiPAP as well
  mutate(device_category = fifelse(is.na(device_category) & 
                                     lag(device_category == "NIPPV") & 
                                     #minute_vent > 1 & ###NEED TO BUILD INTO JHU DATA
                                     pressure_support_set > 1, 
                                   "NIPPV", 
                                   device_category)) |>
  
  mutate(device_category = fifelse(is.na(device_category) & 
                                     lead(device_category == "NIPPV") & 
                                     #minute_vent > 1 & ###NEED TO BUILD INTO JHU DATA
                                     pressure_support_set > 1, 
                                   "NIPPV", 
                                   device_category)) |>
  ungroup()
  
# Now use a Fill Forward Method with Device Category
vent <- vent |>
  group_by(patient_id, encounter_block) |>
  fill(device_category, .direction = 'downup') |>
  ungroup() |>
  as_arrow_table()

#Goal of Function Below is to Define Ventilator Episodes & Ventilator Liberations (> 24 Hours off of MV)
#First will Filter Down to Reduced Table of 'device_category' transitions; For example: This includes rows when a device_category switches from one to another; Also keep First and Last Rows
device_transitions <- vent |> 
  arrange(patient_id, recorded_dttm) |> #Puts in Correct Order
  collect() |>
  group_by(patient_id, encounter_block) |>
  mutate(prev_value_diff = fifelse(
    (device_category!=data.table::shift(device_category, n=1, type = "lag")), 1, 0)) |>
  mutate(prev_value_diff=fifelse(is.na(prev_value_diff), 0, prev_value_diff)) |> #For First Row
  filter(prev_value_diff == 1 |
           row_number()==1 | row_number() == n()) |>
  ungroup()

#Define Ventilator Episodes - Define Ventilator Liberation as 24 Hours Breathing Off Ventilator, Otherwise Will Include That Time in Ventilator Duration
#Temporarily Number Vent Episodes
device_transitions <- device_transitions |>
  group_by(patient_id, encounter_block, device_category) |>
  mutate(category_number=row_number()) |>
  #If a Last Row of an Enconter Block is Not a Device Transition Set Category Number to NA
  group_by(patient_id, encounter_block) |>
  mutate(category_number=fifelse(
    row_number()==n() & prev_value_diff!=1, NaN, category_number
  )) |>
  ungroup() |>
  group_by(patient_id, encounter_block) |>
#Define Vent Start and Stop (Temporary)
  mutate(vent_start=fcase(
    device_category=='IMV' & (prev_value_diff==1 | row_number()==1), recorded_dttm
  )) |>
  mutate(vent_stop=fcase(
    device_category=='IMV' & lead(prev_value_diff)==1, lead(recorded_dttm),
    device_category=='IMV' & lead(row_number())==n(), lead(recorded_dttm) #This is Why we Kept Last Row
  )) |>
  fill(vent_stop, .direction = 'down') |>
  #Define Vent Liberation of Prior Vent Episodes as 24 Hours without device_category=='IMV', can fill backwards for this
  mutate(prior_liberation_new_vent=fcase(
    #This says if the next time someone is on a vent > 24 hours after the last time on a vent it will be a new episode
    device_category=='IMV' & category_number==1, 1, #Need to Define This First
    device_category=='IMV' & recorded_dttm>lag(vent_stop)+dhours(24), 1,
    device_category=='IMV' & recorded_dttm<=lag(vent_stop)+dhours(24), 0
  )) |>
  #Label if Last Row so Vent Duration Can be Defined by Discharge Time
  mutate(last_row=fifelse(row_number()==n(), 1, 0)) |>
  ungroup()

#For Patients Who Start IMV on Last Row can Calculate Time on Vent Using Final Discharge Time
dc_time <- clif_hospitalization |>
  select(patient_id, encounter_block, final_discharge_dttm) |>
  distinct()  |> #1 Row for 1 Encounter Block 
  collect()

device_transitions <- device_transitions |>
  join(dc_time, how = 'left') |>
  #Alternative Way of Labelling Liberation
  group_by(patient_id, encounter_block) |>
  mutate(vent_stop=fifelse(
    device_category=='IMV' & is.na(vent_stop) & last_row==1, final_discharge_dttm, vent_stop)) |>
  mutate(liberation=fcase(
    device_category!='IMV' & recorded_dttm>vent_stop+dhours(24), 1,
    device_category=='IMV' & last_row==1, 0
  )) |>
  fill(liberation, .direction = 'up') |>
  ungroup()
rm(dc_time)

#Renumber 'New' Episodes of MV, that is if the first episode, and then episodes in which the patient was previously liberated, keep the intervening episodes so we can count final duration
vent_episodes <- device_transitions |>
  filter(device_category=='IMV') |>
  group_by(patient_id, encounter_block, prior_liberation_new_vent) |>
  mutate(vent_episode_number=fifelse(
    prior_liberation_new_vent==1, row_number(), NaN)) |>
  group_by(patient_id, encounter_block) |>
  fill(vent_episode_number, .direction = 'down') |>
  group_by(patient_id, encounter_block, vent_episode_number) |>
  mutate(vent_episode_start=fcase(
    row_number()==1, vent_start
  )) |>
  mutate(vent_episode_end=fcase(
    row_number()==n(), vent_stop
  )) |>
  mutate(liberation=fcase(
    row_number()==n(), liberation,
    default = NaN
  )) |>
  fill(vent_episode_start, vent_episode_end, liberation, .direction = 'downup') |>
  #Now Keep First Row for Each Vent Episode
  filter(row_number()==1) |>
  ungroup() |>
  #Calculate Vent Duration
  mutate(vent_duration_hours=as.duration(vent_episode_end-vent_episode_start)/dhours(1)) |>
  select(patient_id, hospitalization_id, encounter_block, device_category, mode_category, 
         liberation:vent_duration_hours)

#Describe Numbers
cat('\nAt this stage in', site, 'data there are', dim(vent_episodes)[1], 'ventilator episodes among', 
    length(unique(vent_episodes$hospitalization_id)), 'hospitalizations from', 
    length(unique(vent_episodes$patient_id)), 'patients.')

#Now Filter to Just First Vent Episodes that are Greater than 36 Hours Long
vent_eligible <- vent_episodes |>
  filter(vent_episode_number==1 & vent_duration_hours>36)

#Describe Numbers
cat(site, '\n After filtering to patients with >36 hours of MV for First Episode there are,', dim(vent_eligible)[1], 'ventilator episodes among,',length(unique(vent_eligible$hospitalization_id)), 'hospitalizations from',length(unique(vent_eligible$patient_id)), 'patients.')

rm(device_transitions)
```
#EXLUSIONS:
-Tracheostomy within 24 hours of first vent
-Patient admitted from OSH on the vent
-After Final Identiication of ABG criteria will then randomly select 1 eligible encounter for each patient with more than 1 eligible
```{r Implement Exclusion Criteria}
#List of patient_id, encounter_block, hospitalization_id and vent_start
temp_ids <- vent_eligible |>
  select(patient_id, encounter_block)

#Tracheostomy in First 24 Hours of First Vent
trach <- clif_respiratory_support |>
  left_join(temp_ids) |>
  filter(!is.na(encounter_block)) |> #THis allows us to keep all encounter block info
  select(patient_id, recorded_dttm, encounter_block, tracheostomy) |>
  filter(tracheostomy==1) |>
  arrange(patient_id, encounter_block, recorded_dttm) |>
  collect() |>
  group_by(patient_id, encounter_block) |>
  filter(row_number()==1) |>
  mutate(first_trach_time=fcase(
    row_number()==1, recorded_dttm
  )) |>
  ungroup() |>
  select(patient_id, encounter_block, tracheostomy, first_trach_time)

#Merge with Vent Eligible and Exclude if first_trach_time within 24 hours of vent start
vent_eligible <- vent_eligible |>
  join(trach, how = 'left') |>
  mutate(tracheostomy=fifelse(is.na(tracheostomy), 0, tracheostomy)) |> #If not merged indicates no trach performed
  mutate(trach_within_24=fcase(
    vent_episode_start+dhours(24)>first_trach_time, 1,
    default = 0
  ))

#Describe for Consort
cat('\n At', site,',', length(unique(vent_eligible$hospitalization_id[vent_eligible$tracheostomy==1])), 'patient hospitalizations were ventilated via a tracheostomy,', length(unique(vent_eligible$hospitalization_id[vent_eligible$trach_within_24==1])), 'within 24 hours of ventilator start and are thus excluded. \n')

#Vent Eligible Excluding Early Trach
vent_eligible <- vent_eligible |>
  filter(trach_within_24!=1)
rm(trach)

#How Many Patients Arrive First to ICU and First Device is a Vent?
#First Define What the First Location Is
osh_transfer <- clif_adt |>
  left_join(temp_ids) |> #Here need to join first and then filter out if encounter ID is missing
  filter(!is.na(encounter_block)) |>
  arrange(patient_id, encounter_block, in_dttm) |>
  collect() |>
  group_by(patient_id, encounter_block) |> # Replace group_column with the column(s) you want to group by
  filter(row_number()==1) |>
  ungroup() |>
  mutate(first_location=location_category)

#Now Define First Device Category
first_device <- clif_respiratory_support |>
  left_join(temp_ids) |>
  filter(!is.na(encounter_block)) |>
  filter(!is.na(device_category)) |>
  arrange(patient_id, encounter_block, recorded_dttm) |>
  collect() |>
  group_by(patient_id, encounter_block) |>
  filter(row_number()==1) |>
  mutate(first_device=device_category) |>
  select(patient_id, encounter_block, first_device)

#Merge back with OSH
osh_transfer <- osh_transfer |>
  join(first_device, how ='left') |>
  mutate(transfer_on_vent=fifelse(
    first_location=='ICU' & first_device=='IMV', 1, 0
  )) |>
  select(patient_id, encounter_block, transfer_on_vent)

#Merge Into Vent Eligible
vent_eligible <- vent_eligible |>
  left_join(osh_transfer) 

#Describe for Consort
cat('\n At', site,',', length(unique(vent_eligible$hospitalization_id[vent_eligible$transfer_on_vent==1])), 'patient hospitalizations were from patients who met criteria for having been transfered while on a ventilator and are thus excluded,') 

#Current Vent Eligible Cohort
vent_eligible <- vent_eligible |>
  filter(transfer_on_vent!=1)
```


